\chapter{Curvature and Topology I}
In this chapter, we will disucss some topological results derived by curvature restrictions.
We will follow the order of Ricci curvature with positive lower bound, nonpositive sectional curvature and constant sectional curvature.
The corresponding results are Bonnet--Myers Theorem, Cartan--Hadamard Theorem and some discussion of distance functions.
More results on topology by curvature restrictions involves comparison theorem, which will be discussed later.

From this chapter, we will assume that all Riemannian manifolds are complete.

\section{Bonnet--Myers Theorem}

Bonnet--Myers Theorem asserts that a Riemannian manifold with positive lower bound for Ricci curvature is bounded, hence compact.
Precisely, we have
\begin{thm}[Bonnet--Myers]
    Let $(M^n,g)$ be a complete Riemannian manifold with $\Ric\geq(n-1)Kg>0$ (here we lowered the index of Ricci tensor), then $\operatorname{diam}(M)\geq\pi/\sqrt{K}$.
    In particular, $M$ is compact.
\end{thm}
\begin{proof}
    We may scale the metric to make $K=1$.
    Use contradiction argument, we may assume there exists $p,q\in M$ with $d(p,q)>\pi$.
    By completeness, there exists a unit speed geodesic $\gamma:[0,a]\to M$ such that $\gamma(0)=p$, $\gamma(a)=q$, so $d(p,q)=a>\pi$, $\gamma$ is length-minimizing.
    Let $\{E_i(t)\}$ be a parallel orthonormal frame along $\gamma$ with $E_n(t)=\dot\gamma(t)$, define
    \[U_i(t)=\sin\left(\frac{\pi t}{a}\right).\]
    Let $I$ be the index form on $\gamma$, then we have
    \begin{align*}
        \sum_{i=1}^{n-1}I(U_i,U_i)&=\sum_{i=1}^n\int_0^a\langle\dot U_i,\dot U_i\rangle-\langle R(\dot\gamma,U_i)\dot\gamma,U_i\rangle\d{t}\\
        &=\sum_{i=1}^{n-1}\int_0^a-\langle\ddot{U}_i,U_i\rangle-\langle R(U_i,\dot\gamma)U_i,\dot\gamma\rangle\d{t}\\
        &=\sum_{i=1}^{n-1}\int_0^a\frac{\pi^2}{a^2}\sin^2\left(\frac{\pi t}{a}\right)-\sin^2\left(\frac{\pi t}{a}\right)\langle R(E_i,\dot\gamma)E_i,\dot\gamma\rangle\d{t}\\
        &=\int_0^a\left((n-1)\frac{\pi^2}{a^2}-\langle\Ric(\gamma),\gamma\rangle\right)\sin^2\left(\frac{\pi t}{a}\right)\\
        &\leq\int_0^a(n-1)\left(\frac{\pi^2}{a^2}-1\right)\sin^2\left(\frac{\pi t}{a}\right)\d{t}\\
        &<0,
    \end{align*}
    the second equality is integration by parts.
    But $\gamma$ is a length-minimizing geo\-desic, $I$ must be positive definite, hence
    \[\sum_{i=1}^{n-1}I(U_i,U_i)\geq 0,\]
    this is a contradiction.
\end{proof}

\begin{rem}
    \begin{enumerate}[(1)]
        \item If we know a priori Cheng's maximal diameter theorem, the above proof can be regarded as a comparison between Jacobi fields on $M$ and sphere.
        Thus this is a prototype of our proof to Rauch's comparison theorem.
        \item The condition cannot be weakened to $K=0$, in fact, $\Sect>0$ is not enough.
        The surface $z=x^2+y^2$ in $\mathbb{R}^3$ is an counterexample.
    \end{enumerate}
\end{rem}

\begin{cor}
    Under above settings, the universal covering $\pi:\tilde{M}\to M$ is compact.
    Moreover, $\pi_1(M)$ is finite.
\end{cor}
\begin{proof}
    We can lift the metric of $M$ to make $\pi$ a Riemannian covering, then $\pi^*g$ has also a Ricci lower bound, and we can apply Bonnet--Myers Theorem.
    For the next claim, let $p\in M$, then $\pi^{-1}(p)$ is a discrete closed set in $\tilde{M}$, hence must be finite since $\tilde{M}$ is compact.
    Then the covering map is of finite sheet, $\pi_1(\tilde{M})$ has finite index in $\pi_1(M)$.
    But $\pi_1(\tilde{M})$ is trivial, hence $\pi_1(M)$ is finite.
\end{proof}

\section{Cartan--Hadamard Theorem}

Cartan--Hadamard Theorem gives the topological characterization of a Riemannian manifold with nonpositive sectional curvature.

\begin{thm}[Cartan--Hadamard]
    Let $M^n$ be a complete Riemannian manifold with $\Sect\leq 0$, then for any $p\in M$ the exponential map $\exp_p:T_pM\to M$ is a universal covering.
\end{thm}

We decompose the proof into several lemmas.

\begin{lem}[Ambrose]\label{Ambrose}
    Let $\varphi:(M,g)\to(N,h)$ be a local isometry.
    If $(M,g)$ is complete, then $\varphi$ is a Riemannian covering map.
\end{lem}
\begin{proof}
    We first show that $\varphi$ is surjective.
    Since $\varphi$ is a local isometry, $\varphi(M)$ is open in $N$.
    Let $y\in N$ be a limit point of $\varphi(M)$, then there exists a $x\in M$ such that there is a geodesic $\sigma:[0,1]\to N$ connecting $\varphi(x)$ and $y$.
    Since $M$ is complete, this geodesic can be lifted to $\tilde\sigma:[0,1]\to M$, and clearly $\varphi(\tilde\gamma(1))=y$.
    Hence $\varphi$ is surjective.

    Now let $x\in M,x'\in N$ with $\varphi(x)=x'$.
    Choose $\delta$ so small that $\exp_{x'}:B'(\delta)\to B_{x'}(\delta)$ is a diffeomorphism, where $B'(\delta)=\{v\in T_{x'}N:\ |v|<\delta\}$.
    Since $\varphi$ is a local isometry, $\varphi^{-1}(x')$ is discrete, say $\{x_i\}_{i\in I}$.
    For each $i\in I$, define
    \begin{gather*}
        B^i(\delta):=\{v\in T_{x_i}M:\ |v|<\delta\},\\
        B^i_\delta:=\{y\in M:\ d(y,x_i)<\delta\}.
    \end{gather*}
    We now show the following three claims sequentially:
    \begin{enumerate}[(1)]
        \item $\varphi^{-1}(B_{x'}(\delta))=\bigcup_{i\in I}B^i_\delta$;
        \item For each $i\in I$, $\varphi:B^i_\delta\to B_{x'}(\delta)$ is a diffeomorphism;
        \item If $i\neq j$, then $B^i_\delta$ and $B^j_\delta$ are disjoint.
    \end{enumerate}

    For (1), $\bigcup_{i\in I}B^i_\delta\subset\varphi^{-1}(B_{x'}(\delta))$ is clear.
    Suppose $z\in\varphi^{-1}(B_{x'}(\delta))$, then there exists a unique geodesic $\gamma:[0,1]\to N$ contained in $B_{x'}(\delta)$ such that $\gamma(0)=x'$, $\gamma(1)=\varphi(z)$ (combine Proposition~\ref{geodesic locally length-min}~and~Corollary~\ref{join points}).
    Since $M$ is complete, we can lift $\gamma$ to $\tilde\gamma$.
    Then by $\varphi(\tilde\gamma(0))=x'$, we have some $\tilde\gamma(0)=x_i$ for some $i\in I$.
    Moreover, since $L[\tilde\gamma]=L[\gamma]<\delta$, we have $z\in B^i_\delta$.
    Hence $\varphi^{-1}(B_{x'}(\delta))\subset\bigcup_{i\in I}B^i_\delta$.
    This is (1).

    For (2), since $M$ is complete, $\exp_{x_i}:B^i(\delta)\to B^i_\delta$ is well-defined.
    By naturality of exponential map (Proposition~\ref{exp natural}), we have the following commutative diagram:
    \[\begin{tikzcd}
        B^i(\delta) \ar[r, "{\d\varphi}"] \ar[d, "{\exp_{x_i}}"] & B'(\delta) \ar[d, "{\exp_{x'}}"] \\
        B^i_\delta \ar[r, "\varphi"] & B_{x'}(\delta).
    \end{tikzcd}\]
    Then
    \[\varphi\circ\exp_{x_i}=\exp_{x'}\circ\d\varphi\]
    is a diffeomorphism.
    Since $\varphi$ is a local isometry, $\exp_{x_i}$ is an immersion, but $\exp_{x_i}$ is surjective, hence $\exp_{x_i}$ is a diffeomorphism.
    Therefore $\varphi=\exp_{x'}\circ\d\varphi\circ(\exp_{x_i})^{-1}$ is a diffeomorphism.
    This is (2).

    For (3), let $z\in B^i_\delta\cap B^j_\delta$ for some $i,j$, then let $\gamma_i:[0,1]\to M$ and $\gamma_j:[0,1]\to M$ be the unique geodesics contained in $B^i_\delta$ and $B^j_\delta$ with $\gamma_i(0)=\gamma_j(0)=z$, $\gamma_i(1)=x_i$ and $\gamma_j(1)=x_j$.
    Let $\zeta:[0,1]\to N$ be the unique geodesic contained in $B_{x'}(\delta)$ with $\zeta(0)=\varphi(z)$, $\zeta(1)=x'$.
    Then by uniqueness, we have $\varphi\circ\gamma_i=\varphi\circ\gamma_j$.
    However, since $\varphi$ is a local isometry, we have
    \[\dot\gamma_i(0)=\d\varphi^{-1}(\dot\zeta(0))=\dot\gamma_j(0)\]
    and
    \[L[\gamma_i]=L[\zeta]=L[\gamma_j],\]
    hence $\gamma_i(1)=\gamma_j(1)$, that is, $x_i=x_j$.
    This is (3).
\end{proof}

\begin{lem}\label{C-H 1}
    Let $(M,g)$ be a Riemannian manifold, $p\in M$.
    If $M$ contains no conjugate points of $p$, then $\exp_p:T_pM\to M$ is a covering map.
\end{lem}
\begin{proof}
    Since $p$ has no conjugate points, $\exp_p$ is nonsingular on $T_pM$, hence is an immersion.
    Then we can equip $T_pM$ with pullback metric $\exp_p^*g$.
    If we can show $(T_pM,\exp_p^*g)$ is complete, then by Lemma~\ref{Ambrose}, $\exp_p$ is a covering map.
    Let $\gamma:[0,a)\to T_pM$ be a length-minimizing geodesic, then since $\exp_p$ is a local isometry between $(T_pM,\exp_p^*g)$ and $(M,g)$, $\exp_p\circ\gamma:[0,a)\to M$ must be a geodesic.
    By the definition of exponential map, we have $\exp_p\circ\gamma(t)=\exp_p(\dot\gamma(0)t)$, then it clearly can be extended to $a$.
    Thus by Theorem~\ref{Hopf-Rinow}~(4), $(T_pM,\exp_p^*g)$ is complete.
    This proves the lemma.
\end{proof}

\begin{lem}\label{C-H 2}
    Let $(M,g)$ be a complete Riemannian manifold with $\Sect\leq 0$, then for any $p\in M$, $p$ contains no conjugate points.
\end{lem}
\begin{proof}
    Only need to show for any geodesic $\gamma:[0,+\infty)\to M$ with $\gamma(0)=p$, the nontrivial Jacobi field $J$ along $\gamma$ with $J(0)=0$ has no zeros besides $0$.
    Consider the function $f(t)=\langle J(t),J(t)\rangle$, then we have
    \[\dot{f}(t)=2\langle J(t),\dot{J}(t)\rangle,\]
    and
    \begin{align*}
        \ddot{f}(t)&=2(\langle\dot{J}(t),\dot{J}(t)\rangle+\langle J(t),\ddot{J}(t)\rangle)\\
        &=2(\langle\dot{J}(t),\dot{J}(t)\rangle-\langle R(\dot\gamma,J)\dot\gamma,J\rangle(t)\\
        &\geq\langle\dot{J}(t),\dot{J}(t)\rangle.
    \end{align*}
    Hence
    \begin{align*}
        \dot{f}(t)&\geq\int_{0}^{t}\langle\dot{J}(t),\dot{J}(t)\rangle\d{t}+\dot{f}(0)\\
        &=\int_{0}^{t}\langle\dot{J}(t),\dot{J}(t)\rangle\d{t}+\langle\dot{J}(0),J(0)\rangle\\
        &=\int_{0}^{t}\langle\dot{J}(t),\dot{J}(t)\rangle\d{t}\\
        &>0,
    \end{align*}
    unless $\dot{J}(t)\equiv 0$ for all $t\geq 0$, but this contradicts $J$ being nontrivial.
    Hence by integrating $\dot{f}(t)$ we obtain $|J(t)|^2>0$ for $t>0$, i.e.\ $J$ has no zero other than $0$.
\end{proof}

Now Cartan--Hadamard Theorem is the corollary of Lemma~\ref{C-H 1}~and~\ref{C-H 2}.


\section{Cut Loci and Distance Functions}

Distance functions control the topology of a Riemannian manifold via comparison theorems.
But before we discuss comparison theorems, we first need to go through distance functions in detail.

We start this section by an example.

\begin{eg}
    Consider the cylinder $M=\mathbb{S}^1\times\mathbb{R}$ with product metric.
    Then $M$ is flat, and by Cartan--Hadamard Theorem, it has no conjugate points.
    Let $\gamma:[0,2\pi]\to M$ be a generatrix circle, then $\gamma$ is a geodesic.
    Denote $p=\gamma(0)$, $q=\gamma(\pi)$, then $\gamma|_{[0,\pi]}$ is a length-minimizing geodesic joining $p$ and $q$.
    But it is not the only length-minimizing geodesic joining $p$ and $q$, as $-\gamma|_{[\pi,2\pi]}$ is another one.
    Then $q$ spoils the uniqueness of length-minimizing geodesic starts from $p$, without being conjugate to $p$.
    This inspires us to denote the notion of cut points.
\end{eg}

Given $v\in T_pM$, $|v|=1$, $\gamma_v:[0,+\infty)\to M$ be a geodesic with $\gamma_v(0)=p$, $\dot\gamma_v(0)=v$.
Notice that $\gamma_v|_{[0,t_0]}$ is length-minimizing if $d(p,\gamma_v(t_0))=t_0$.

\begin{defn}
    Under above settings, let $t_0=\sup\{t\in(0,+\infty):\ d(p,\gamma_v(t))=t\}$.
    If $t_0<+\infty$, we call $\gamma_v(t_0)$ the \emph{cut point} of $p$ along $\gamma_v$.
    The \emph{cut locus} of $p$ is the set of all cut points of $p$, denoted by $\Cut{p}$.
\end{defn}

By the definition and Jacobi Theorem (Theorem~\ref{jacobi thm}), the first conjugate point (if exists) must be cut point.
But converse is generally not true, see Example~\ref{eg of cut loci}.
The following proposition characterizes cut points.

\begin{prop}
    Let $\gamma$ be a unit speed geodesic starts from $p$.
    If $\gamma(t_0)$ is the cut point of $p$ along $\gamma$, then either $\gamma(t_0)$ is conjugate to $p$, or there exists two length-minimizing geodesics from $p$ to $\gamma(t_0)$.
\end{prop}
\begin{proof}
    Let the initial vector of $\gamma$ be $v$.
    Choose a sequence $\{t_i\}$ decreasingly converges to $t_0$.
    Let $\sigma_i$ be the unit speed length-minimizing geodesic joining $p$ and $t_i$, with initial vector $v_i$.
    Define $b_i=d(p,\sigma_i(t_i))=L[\sigma_i]$, then $\sigma_i$ is defined on $[0,b_i]$ and $b_i\to t_0$.
    Hence up to a subsequence, we have $v_i\to y$ for some $y\in T_pM$ with $|y|=1$.
    Now we have two cases.
    \begin{enumerate}[(1)]
        \item If $v\neq y$, then by some ODE theory, $\sigma_i\to\sigma(t)=\exp_p(ty)$ (up to a subsequence).
        By the lower semi-coninuity of length functional, we have
        \[L[\sigma]\leq\lim_{i\to\infty}L[\sigma_i]\leq\lim_{i\to\infty}t_i=t_0.\]
        On the other hand, $\gamma|_{[0,t_0]}$ is length-minimizing, we have
        \[L[\sigma]\geq L[\gamma|_{0,t_0}]=t_0.\]
        Hence $L[\sigma]=t_0$, $\sigma$ is another length-minimizing geodesic joining $p$ and $\gamma(t_0)$.
        \item If $v=y$, then
        \[\exp_p(b_iv_i)=\gamma(t_i)=\exp_p(t_iv)\]
        with both $b_iv_i\to t_0v$ and $t_iv\to t_0v$.
        This means $\exp_p$ is not injective in a neighborhood of $\gamma(t_0)$, hence $\exp_{p*}|_{\gamma(t_0)}$ degenerates, $\gamma(t_0)$ is conjugate to $p$.
    \end{enumerate}
\end{proof}

\begin{eg}\label{eg of cut loci}
    \begin{enumerate}[(1)]
        \item Let $M=\mathbb{S}^n$.
        Then the south pole is conjugate to north pole, as well as the cut point of north pole.
        \item Let $M=\mathbb{RP}^n$.
        For simplicity we take $n=2$.
        We regard $\mathbb{RP}^2$ as the upper hemisphere with antipodal points on equator identified.
        Then from north pole we move along a ``great circle'' to a point $p$ on equator, the path is a length-minimizing geodesic.
        However, the path on the same ``great circle'' but on the other side is also a length-minimizing geodesic, so $p$ is the cut point of north pole (along this geodesic).
        But by the naturality of exponential map, $p$ is not conjugate to north pole, so this is an example of cut point not being conjugate point.
        \item Let $M=\mathbb{R}^2/\mathbb{Z}^2$ be the flat torus.
        Let $[0,1]\times[0,1]$ be the fundamental region, and denote $p=(0,0)=(0,1)$.
        Then $\gamma(t)=(t,0)$ is a geodesic.
        Notice that if $t_0>1/2$, $\gamma(t)$ is not the length-minimizing geodesic joining $p$ and $(t_0,0)$, as $\sigma(t)=(1-t,0)$ is.
        By the same reason, we have the cut locus of $p$ is $\{1/2\}\times[0,1/2]\cup[0,1/2]\times\{1/2\}$.
    \end{enumerate}
\end{eg}

Let $S_p\subset T_pM$ be the unit sphere.
For any $v\in S_p$, denote $\gamma_v:[0,+\infty)\to M$ be the geodesic with $\dot\gamma(0)=v$.
We define a function
\[\tau:S_p\to(0,+\infty],\ \tau(v)=\begin{cases}
    +\infty, & \gamma_v\text{ contains no cut point},\\
    t_0, & \gamma_v(t_0)\text{ is the cut point of }p.
\end{cases}\]

\begin{prop}
    $\tau$ is continuous.
\end{prop}
\begin{proof}
    I don't know!
\end{proof}